
#Projecr Group 12
#Members : Ashwin , Zach , Upendra , Nita


#About the DataSet :
#Bike sharing systems are new generation of traditional bike rentals where whole process from membership, rental and return back has become automatic. 
#Through these systems, user is able to easily rent a bike from a particular position and return back at another position. 
#Currently, there are about over 500 bike-sharing programs around the world which is composed of over 500 thousands bicycles. Today, there exists great interest in these systems due to their important role in traffic, environmental and health issues. 

#The following graphs shown below is attempt to show descriptive analytics of bike sharing dataset.
#We have try to visualize the key feature which impact the business of the bike sharing company.



```{r}
#Loading the libraries 
library(ggplot2)
library(dplyr)
library(tidyr)
library(stringr)
library(ggplot2)
library(gridExtra)
# library(lubridate)

```



```{r}
#Loading the data

df_day <- read.csv("day.csv")
df_hour <- read.csv("hour.csv")

```



```{r}

#Plot 1: Density Plot

ggplot(df_hour) +
 aes(x = cnt) +
 geom_histogram(bins = 97L, fill = "#112446") +
 labs(x = "Number of Bikes Rented Per Hour", 
 y = "Frequency", title = "Density of Bike Rentals") +
 theme_bw()

summary(df_hour$cnt)


#The below graph show the density of Bikes rented per hour.
#It shows a skewed distribution. The number of bikes rented per hour are less than 200 and the max is 977


```

```{r}
#Plot 2: Boxplot

ggplot(df_day,aes(factor(yr),cnt)) +
  geom_boxplot(shape = "circle", fill = "#0B4B89") +
  labs(x = "Years", y = "Count") +
  labs(title = "Summary of rental bikes per year") +
  scale_x_discrete(labels = c("2011","2012"))

#The below boxplot is for each year which show the summary and also the outlier present in the data.

```





```{r ,fig.width=10}

#Plot 3 : Dodge Bar Plot

df_day$season[df_day$season==1] <- "Spring"
df_day$season[df_day$season==2] <- "Summer"
df_day$season[df_day$season==3] <- "Fall"
df_day$season[df_day$season==4] <- "Winter"

df_day$weathersit[df_day$weathersit==1] <- "Clear, Few clouds, Partly cloudy"
df_day$weathersit[df_day$weathersit==2] <- "Mist + Cloudy, Broken clouds, Few clouds,"
df_day$weathersit[df_day$weathersit==3] <- "Light Snow, Light Rain + Thunderstorm + Scattered clouds,"
df_day$weathersit[df_day$weathersit==4] <- "Heavy Rain + Ice Pallets + Thunderstorm + Mist, Snow + Fog"

## Count of rental, if its casual or registered user
result1 <- df_day %>% group_by(mnth) %>%
  summarise(Average_Casual_Users = mean(casual),
            Average_registered_Users= mean(registered)) 

result_pivot <- result1 %>% pivot_longer(2:3, names_to =  "Type of User", values_to = "Number of User")

nb<-c("Jan",  "Feb","Mar","Apr","May","Jun","Jul","Aug","Sep","Oct","Nov","Dec")

names(result_pivot)[1] <- "Month"

class(result_pivot$Month)<-"numeric"

ggplot(data = result_pivot, aes(x=round(Month), y=round(`Number of User`)))+ 
  geom_bar(stat= "identity",aes(fill= `Type of User`),position = "dodge", width = 0.9)+
  geom_text(aes(label=round(`Number of User`),group=`Type of User`),position= position_dodge(width = 1), color="black",   size=3.5, vjust= -0.5,) +scale_x_continuous(breaks=seq(1,12,1),labels=nb)+ ylim(0,5000) +
  ggtitle("Casual and Registered User comparison per Month") +
  xlab("Month")+
  ylab("Number of Users")+
  theme(plot.title = element_text(face="bold",color = "darkblue", size = 20) ) +
  theme(legend.title=element_text(face="bold"))+theme(axis.text.x = element_text(face="bold", color="black", 
                           size=14, ),
          axis.text.y = element_text(face="bold", color="#993333", 
                           size=14, angle=45))

#Comparision between Registered and Casual user for each month.

```


```{r}

#Plot 4: Stack bar plot


result2 <- df_day %>% group_by(season) %>%
  summarise(Average_Casual_Users = mean(casual),
            Average_registered_Users= mean(registered)) 
result_pivot1 <- result2 %>% pivot_longer(2:3, names_to =  "Type of User", values_to = "Number of User")


 ggplot(data = result_pivot1, aes(x= reorder(season,+`Number of User`), y=round(`Number of User`)))+ 
  geom_bar(stat= "identity",aes(fill= `Type of User`), width = 0.9,na.rm=TRUE,position = "stack")+
  geom_text(aes(label=round(`Number of User`),group=`Type of User`), color="black", size=6, hjust= 1,position ="stack",)  + coord_flip()+scale_fill_manual(values = c("orange","skyblue"))+
  ggtitle("Casual and Registered User comparison per Season") +
  xlab("Season")+
  ylab("Number of Users")+
  theme(plot.title = element_text(face="bold",color = "darkblue", size = 20) ) +
  theme(legend.title=element_text(face="bold"))+theme(axis.text.x = element_text(face="bold", color="#993333", 
                           size=14 ),
          axis.text.y = element_text(face="bold", color="black", 
                           size=14 ))

#Seasion wise comparison of Casual and Registered user
# Below graph gives us the idea of the average number of users based on the seasons who likes to casually ride the bike compared users who are registered
 
 
```




```{r,fig.width= 10}
#Plot 5: Heatmap/Correlation Plot

ggplot(df_day, aes(x=weathersit, y=season)) +
  geom_tile(aes(fill = cnt), colour = "white") +
  scale_fill_gradient("Count of Users",low = "white", high = "Red")+ theme(plot.title = element_text(face="bold",color = "darkblue", size = 20) ) +
  theme(legend.title=element_text(face="bold", size=20)) +theme(axis.text.x = element_text(face="bold", color="#993333", 
                           size=14, ),
          axis.text.y = element_text(face="bold", color="#993333", 
                           size=14, angle=45)) +xlab("Weather Conditions") +ylab("Seasons")+ggtitle("Heat Map ")


## when the weather is clear , partly cloudy people like to ride the bikes, because there is no rain or snow or thunderstorm
#The heat map below displays the relation between total bike rentals as a function of season based on weather conditions. During the most ideal weather days, the total bike rentals are the highest followed closely by the slightly cloudy weather. when the weather is the worst leading to a sharp decrease in total bikes rented.
```




```{r}
#Plot 6: : Line Graph

ch = aggregate(cnt~hr, mean, data = df_hour)
plot(ch$hr, ch$cnt, type = "b", col = "#33658A", lwd = 2.5, xlab = "hour of the day",
     ylab = "Average no. of bike rentals", main = "Bike rentals vs. hour of the day", 
     cex.main = 2, cex.lab = 1.5, cex.axis = 1.3, cex = 1.2)

#count of bicylces with hour of the day
#The Peaks hours are show below 

```



```{r}

#Plot 7: Pie Chart
# class(df_day$workingday)="logical"
# test = df_day %>%
#   aggregate(by = list(df_day$workingday), FUN = mean)%>%
#   select(c("workingday","cnt"))

test <- df_day %>% group_by(workingday) %>%
  summarise(cnt= mean(cnt))

test$workingday[test$workingday == FALSE] = "Working Days"
test$workingday[test$workingday == TRUE] = "Non-Working/Holiday"

data <- test %>% 
  mutate(prop = cnt / sum(test$cnt) *100) %>%
  mutate(ypos = cumsum(prop)- 0.5*prop )

ggplot(data, aes(x = "", y = prop, fill = workingday)) +
  geom_col(color = "black") +
  geom_label(aes(label = c("49 %","51 %")), color = "white",
            position = position_stack(vjust = 0.5),
            show.legend = FALSE) +
  guides(fill = guide_legend(title = "Answer")) +
  coord_polar(theta = "y") + 
  ggtitle("                 Proportion of Rental Bikes on Working and Non-Working days") +
  theme_void()

#Rentals Proportion on holidays vs working days
#There is not much difference in the proportion


```


```{r}
#Plot 8 : Scatter Plot

# convert feel temperature to general scale
df_hour$feel.temp = df_hour$atemp*(50+16) - 16
# summary(df_hour$feel.temp)

scatter.smooth(df_hour$feel.temp,df_hour$cnt, col = "#33658A", xlab = "Feel temperature", ylab = "Number of Bike Rentals", main = "Bike Rentals vs. Feel Temperature", cex.main = 2, cex.lab = 1.5, cex.axis = 1.3)

plot <- ggplot(df_hour, aes(feel.temp, cnt)) + geom_smooth(aes(color = cnt))
plot + xlab("Feel Temperature °C") + ylab("Avg Number of Bike Rentals") +
  theme_light(base_size = 11) + scale_x_continuous(expand = c(0, 0)) + scale_y_continuous(expand = c(0, 0)) +
  theme(text = element_text(size = 20), plot.title = element_text(size = 24, face="bold"))


#Below scatter plot shows the distribution of number of bike rentals against the feel temperature. The feel temperature is a measure of how hot or cold it feels like outside.

#For Bike rentals per hour and the feel temperature, we can assume that people are more likely to ride bikes as the feel temperature gets higher.

```

